---
output:
  pdf_document: default
  html_document: default
---

The Effect of Natural Disasters on Incumbent Support
---
*Authored by Jack Holland, Jenny Lu, and Menemsha Solomon*

Research Question: How does a natural disaster affect candidate performance?
---
___

Experimental Design:
---

**Treatment Variable**: Exposure to Hurricane Sandy (i.e. a certain threshold of property damage as assessed by disaster relief)
  - Threshold will be decided upon looking at hurricane damage data set
  - *Data set*: Sandy Damage Estimates by Block Group

**Outcome Variable**: 2008-2012 Election Trend
  - (2012 election D margin - 2008 election D margin) - (Nationwide D swing [-3.3])

**Constants**: Demographic Data
  - Needs to be disaggregated into block data for merging
    - https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/T9VMJO
    - Also helpful for sourcing data: https://davesredistricting.org/maps#aboutdata \

The first step of this process if to find and aggregate official election results by precinct from both 2008 and 2012. Unfortunately, there exists no centralized data set for either of these elections, so we must acquire them on an individual basis and then clean and merge them.\

#Precinct-Level Election Returns
```{r}
setwd("/Users/jackholland/Downloads/School/GOV51/Final Project")

#NEW YORK
ny_2008 <- read.csv("NY_2008.csv")
ny_2012 <- read.csv("NY_2012.csv")

#CONNECTICUT
ct_2008 <- read.csv("CT_2008.csv")
ct_2012 <- read.csv("CT_2012.csv")

#NEW JERSEY
```

#Demographic data by block
```{r}
setwd("/Users/jackholland/Downloads/School/GOV51/Final Project")

#NEW YORK

#CONNECTICUT
ct_cvap_2010 <- read.csv("ct_cvap_2010.csv")


#NEW JERSEY
```

