---
output:
  pdf_document: default
  html_document: default
---

The Effect of Natural Disasters on Incumbent Support
---
*Authored by Jack Holland, Jenny Le, and Menemsha Solomon*

Research Question: How does a natural disaster affect candidate performance?
---
___

Experimental Design:
---

**Treatment Variable**: Exposure to Hurricane Sandy (i.e. a certain threshold of property damage as assessed by disaster relief)
  - Threshold will be decided upon looking at hurricane damage data set
  - *Data set*: Sandy Damage Estimates by Block Group

**Outcome Variable**: 2008-2012 Election Trend
  - (2012 election D margin - 2008 election D margin) - (Nationwide D swing [-3.3])

**Constants**: Demographic Data
  - Needs to be disaggregated into block data for merging
    - https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/T9VMJO
    - Also helpful for sourcing data: https://davesredistricting.org/maps#aboutdata \

The first step of this process is to find and aggregate official election results by precinct from both 2008 and 2012. Unfortunately, there exists no centralized data set for either of these elections, so we must acquire them on an individual basis and then clean and merge them.

#Precinct-Level Election Returns
```{r}
#CONNECTICUT
ct_2008 <- read.csv("20081104__ct__general__town.csv")
ct_2012 <- read.csv("20121106__ct__general__precinct.csv")

##Cleaning 2008 data
ct_2008 <- subset(ct_2008, office == "President")
ct_2008 <- subset(ct_2008, candidate == "Barack Obama" | candidate == "John McCain")
ct_2008 <- subset(ct_2008, county != "Total")
ct_2008 <- ct_2008[, c(2, 3, 5, 6, 7)]
ct_2008$party <- ifelse(ct_2008$party == "Dem", "DEM", "REP")

##Cleaning 2012 data
ct_2012 <- subset(ct_2012, office == "President")
ct_2012 <- subset(ct_2012, candidate == "Barack Obama" | candidate == "Mitt Romney")
ct_2012 <- aggregate(votes ~ town + candidate + party + office, data = ct_2012, sum)

##Cleaning original data so the merge is seamless
ct_2012$town <- ifelse(ct_2012$town == "BeaconFalls", "Beacon Falls", ct_2012$town)
ct_2012$town <- ifelse(ct_2012$town == "DeepRiver", "Deep River", ct_2012$town)
ct_2012$town <- ifelse(ct_2012$town == "EastHaddam", "East Haddam", ct_2012$town)
ct_2012$town <- ifelse(ct_2012$town == "EastHampton", "East Hampton", ct_2012$town)
ct_2008$town <- ifelse(ct_2008$town == "Litchfeld", "Litchfield", ct_2008$town)
ct_2012$town <- ifelse(ct_2012$town == "Meridien", "Meriden", ct_2012$town)
ct_2008$town <- ifelse(ct_2008$town == "New Mil!ford", "New Milford", ct_2008$town)
ct_2012$town <- ifelse(ct_2012$town == "NorthCanaan", "North Canaan", ct_2012$town)
ct_2012$town <- ifelse(ct_2012$town == "NorthStonington", "North Stonington", ct_2012$town)
ct_2012$town <- ifelse(ct_2012$town == "OldLyme", "Old Lyme", ct_2012$town)
```

```{r}
##combining rows with same town name
library(tidyverse)
ct2012 <- ct_2012 |>
  group_by(town) |>
  mutate(row_id = row_number()) |>
  pivot_wider(names_from = row_id, values_from = c(candidate, party, office, votes), names_prefix = "Row", values_fill = NULL) |>
  ungroup() 

ct2012 <- ct2012 |>
  rename("2012_dem_candidate" = candidate_Row1,
         "2012_rep_candidate" = candidate_Row2,
         "2012_dem_votes" = votes_Row1,
         "2012_rep_votes" = votes_Row2) |>
  select(town, "2012_dem_candidate", "2012_dem_votes", "2012_rep_candidate", "2012_rep_votes") |>
  mutate(`2012_total` = `2012_dem_votes` + `2012_rep_votes`) |>
  mutate(`2012_repvoteshare` = `2012_rep_votes` / `2012_total`) |>
  mutate(`2012_demvoteshare` = `2012_dem_votes` / `2012_total`) 

ct2008 <- ct_2008 |>
  group_by(town) |>
  mutate(row_id = row_number()) |>
  pivot_wider(names_from = row_id, values_from = c(candidate, party, office, votes), names_prefix = "Row", values_fill = NULL) |>
  ungroup() 

ct2008 <- ct2008 |>
  rename("2008_dem_candidate" = candidate_Row1,
         "2008_rep_candidate" = candidate_Row2,
         "2008_dem_votes" = votes_Row1,
         "2008_rep_votes" = votes_Row2) |>
  select(town, "2008_dem_candidate", "2008_dem_votes", "2008_rep_candidate", "2008_rep_votes") |>
  mutate(`2008_total` = `2008_dem_votes` + `2008_rep_votes`) |>
  mutate(`2008_repvoteshare` = `2008_rep_votes` / `2008_total`) |>
  mutate(`2008_demvoteshare` = `2008_dem_votes` / `2008_total`) 

##Merging
merged_CT <- merge(ct2008, ct2012, by = c("town"), all = TRUE) |>
  mutate(dif_rep_voteshare = `2012_repvoteshare` - `2008_repvoteshare`)|>
  mutate(dif_dem_voteshare = `2012_demvoteshare` - `2008_demvoteshare`) |>
  mutate(rep_vote_percent_diff = (`2012_rep_votes` - `2008_rep_votes`)/`2008_rep_votes`) |>
  mutate(dem_vote_percent_diff = (`2012_dem_votes` - `2008_dem_votes`)/`2008_dem_votes`) |>
  mutate(votedif = (`2012_total` - `2008_total`)/`2008_total`)

##Create a table with only variables of interest, voteshare difference from 2008 to 2012
CT_votes <- merged_CT |>
  select(town, rep_vote_percent_diff, dem_vote_percent_diff, votedif, `2008_total`, `2012_total`)

##List of coastal towns
coastal_towns <- c("Bridgeport", "Darien", "Fairfield", "Greenwich", "Norwalk", "Stamford", 
                   "Stratford", "Westport", "Branford", "East Haven", "Guilford", "Madison", 
                   "Milford", "New Haven", "West Haven", "Clinton", "Old Saybrook", "Westbrook", 
                   "East Lyme", "Groton", "New London", "Stonington", "Waterford")

##Creating column for coastal towns
CT_votes$Location <- ifelse(CT_votes$town %in% coastal_towns, "Coastal", "Non-Coastal")

##Calculating Point Estimate 
point_estimates <- CT_votes %>%
  group_by(Location) %>%
  summarize(
    mean_rep_vote_diff = mean(rep_vote_percent_diff),
    mean_dem_vote_diff = mean(dem_vote_percent_diff),
    se_rep_vote_diff = sd(rep_vote_percent_diff) / sqrt(n()), # Standard error for Republican vote difference
    se_dem_vote_diff = sd(dem_vote_percent_diff) / sqrt(n())  # Standard error for Democratic vote difference
  )
```

```{r}
##Plot for voteshare
ggplot(CT_votes, aes(x = rep_vote_percent_diff, y = dem_vote_percent_diff, label = town)) +
  geom_point(aes(color = Location, alpha = Location), size = 3) +
  scale_color_manual(values = c("Coastal" = "blue", "Non-Coastal" = "orange")) + ##make non-coastal translucent
  xlim(-1.15, 0.6) + ##adjusting axis parameters so dots fit better
  ylim(-0.5, 7) +
  scale_alpha_manual(values = c("Coastal" = 1, "Non-Coastal" = 0.3)) +
  geom_text(aes(label = town), hjust = 1.1, vjust = 1.1, check_overlap = TRUE, size = 3) + ##label some times w/o overlap
  labs(
    x = "Difference in Republican Vote Share",
    y = "Difference in Democratic Vote Share",
    title = "Vote Share Differences in Towns Across Connecticut (2008 to 2012)"
  ) +
  theme(
    plot.title = element_text(size = 13),
    axis.title = element_text(size = 13)
  )
```
```{r}
##Plot for vote cast differences
ggplot(CT_votes, aes(x = factor(1), y = votedif, fill = Location)) +
  geom_jitter(width = 0.1, size = 3, shape = 21, color = "black", alpha = 0.6) +
  scale_fill_manual(values = c("Coastal" = "blue", "Non-Coastal" = "orange")) +
  labs(x = "Difference in Votes Cast", y = "Vote Cast Differences", title = "Vote Cast Difference in Towns Across Connecticut (2008 to 2012)") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  xlab(NULL) # Remove x-axis label

```


```{r}
library(stringr)
library(tidyverse)
library(dplyr)

##load and clean GEOID, ZCTA, Income, and Town data

geoid_zcta5 <- read.csv("CT_ZCTA5_GEOID.csv")

geoid_zcta5 <- geoid_zcta5 |>
  select(ZCTA5, GEOID)

zcta5_town <- read.csv("ZCTA_CT_7124349094171390284.csv")

zcta5_town <- zcta5_town |>
  rename(ZCTA5 = ZCTA5CE10) |>
  select(ZCTA5, Town)

ct_incomes <- read.csv("ct_incomes.csv")

##put census data in terms of town
census <- read.csv("ct_cvap_2010.csv") 
census2 <- census |>
  mutate(GEOID = str_sub(as.character(GEOID), 1, 10)) |>
  mutate(GEOID = as.numeric(GEOID))

census_geoid <- merge(census2, geoid_zcta5, by = "GEOID")

census_zcta5 <- merge(zcta5_town, census_geoid, by = "ZCTA5")

census_zcta5 <- census_zcta5 |>
  select(Town, CVAP_TOT10, CVAP_NHS10, CVAP_AIA10, CVAP_ASN10, CVAP_BLK10, CVAP_NHP10, CVAP_WHT10, CVAP_AIW10,
         CVAP_ASW10, CVAP_BLW10, CVAP_AIB10) 

census_town_data <- aggregate(.~ Town, data = census_zcta5, sum)

##put damage data in terms of town
damage <- read.csv("FemaDamageAssessmnts_01172013_new_8825820061294884462.csv")
damage <- damage |>
  rename(GEOID = X2010.Census.Block.Group.ID) |>
  filter(str_sub(GEOID, 1, 2) == "09") 

damage <- damage |>
  mutate(GEOID = as.numeric(GEOID)) |>
  mutate(GEOID = str_sub(as.character(GEOID), 1, 10)) |>
  mutate(GEOID = as.numeric(GEOID))

damage_geoid <- merge(damage, geoid_zcta5, by = "GEOID")

damage_zcta5 <- merge(zcta5_town, damage_geoid, by = "ZCTA5")

damage_zcta5 <- damage_zcta5 |>
  select(Town, Total.Damaged..any.type., Total.Housing.Units) 

damage_town_data <- aggregate(.~ Town, data = damage_zcta5, sum)

## combine census and damage data 
town_data <- merge(census_town_data, damage_town_data, by = "Town", all = TRUE)

town_data <- town_data |>
  rename(town = Town)

##combine census, damage, and voting data
finaldata <- merge(CT_votes, town_data, by = "town", all = TRUE)
finaldata <- merge(finaldata, ct_incomes, by = "town", all = TRUE)

##cleaning numeric income values
finaldata$Per.capita.income <- as.numeric(gsub("[\\$,]", "", finaldata$Per.capita.income))
finaldata$Median.household.income <- as.numeric(gsub("[\\$,]", "", finaldata$Median.household.income))
finaldata$Population <- gsub(",", "", finaldata$Population, fixed = TRUE)
finaldata$Population <- as.numeric(finaldata$Population)

##row for percentage white non-hispanic, storm cindition column
finaldata <- finaldata |>
  mutate(pct_white = CVAP_WHT10/CVAP_TOT10, na.rm = TRUE) |>
  mutate(Total.Damaged..any.type. = log(Total.Damaged..any.type.)) |>
  mutate(Total.Damaged..any.type. = if_else(is.na(Total.Damaged..any.type.), 0, Total.Damaged..any.type.)) |>
  mutate(condition = if_else(Total.Damaged..any.type. == 0, "No Damage", "Damage")) |>
  mutate(dem_raw_vote_swing = round((dem_vote_percent_diff*Population)*0.043)) |>
  mutate(coastal_binary = if_else(Location == "Coastal", 1, 0))

View(finaldata)
```

#Tests
```{r}
##subset based on damage treatment
damaged_subset <- finaldata[finaldata$Total.Damaged..any.type. != 0, ]
control_subset <- finaldata[finaldata$Total.Damaged..any.type. == 0, ]

##regression
regression_1 <- lm(dem_vote_percent_diff ~ Total.Damaged..any.type. + pct_white + Population + Per.capita.income + Median.household.income + coastal_binary, data = finaldata)
summary(damage_regression)

damaged_subset_regression <- lm(dem_vote_percent_diff ~ Total.Damaged..any.type. + pct_white + Population + Per.capita.income + Median.household.income + coastal_binary, data = damaged_subset)
summary(damaged_subset_regression)

#graphing
outlier_removed <- damaged_subset[damaged_subset$Total.Damaged..any.type. <= 500 & damaged_subset$dem_vote_percent_diff <=2, ]
outlier_removed$dem_vote_percent_diff <- as.numeric(outlier_removed$dem_vote_percent_diff)

damage_viz <- ggplot(data = outlier_removed, aes(x = Total.Damaged..any.type., y = dem_vote_percent_diff)) +
  geom_point(aes(color = dem_vote_percent_diff), size = 3) +
  geom_smooth(method = "lm", color = "blue") +
  geom_text(aes(label = town), vjust = -0.5, size = 3) +
  labs(x = "Number of Properties Damaged", y = "Dem Vote Percent Difference", title = "Property Damage vs. Democratic Vote Shift (Subsetted to Towns Damaged)", color = "Dem. Vote Shift (%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_color_gradient2(low = "red", high = "royalblue", midpoint = 0, guide = "colourbar") +
  geom_hline(yintercept = 0, alpha = 0.5)

damage_viz

raw_swing_map <- ggplot(data = outlier_removed, aes(x = Total.Damaged..any.type., y = dem_raw_vote_swing)) +
  geom_point(aes(color = dem_raw_vote_swing), size = 3) +
  geom_smooth(method = "lm", color = "blue") +
  geom_text(aes(label = town), vjust = -0.5, size = 3) +
  labs(x = "Number of Properties Damaged", y = "Dem Vote Raw Difference", title = "Property Damage vs. Democratic Raw Vote Shift (Subsetted to Towns Damaged)", color = "Dem. Vote Shift") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_color_gradient2(low = "red", high = "royalblue", midpoint = 0, guide = "colourbar") +
  geom_hline(yintercept = 0, alpha = 0.5)

raw_swing_map

##DID
damaged_votes <- mean(finaldata$dem_vote_percent_diff[finaldata$condition == "Damage"], na.rm = TRUE)
no_damage_vote <- mean(finaldata$dem_vote_percent_diff[finaldata$condition == "No Damage"], na.rm = TRUE)

dif_damage_numeric = (damaged_votes - no_damage_vote)
dif_damage_numeric
```

